Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-16 paper/main.tex
results

We use the implementation of GECO provided by the
authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the
authors of CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use and
the GA that optimizes the surrogate models uses a population size of 20,
crossover rate of 0.5, mutation rate of 1, tournament selection of size 2 and
probability of 0.9, interchange mutation (i.e., exchanging two randomly
selected elements of the permutation) and cycle crossover for permutations. The
budget of each run of the GA is \textcolor{red}{$10^4$} evaluations using the
surrogate-model. Although it is never stated in the original paper, the	`
implementation of CEGO generates a set of initial solutions of size
$\minit=10$ by means of a max-min-distance sequential design: new solutions
are added to the set sequentially by maximizing the minimum distance to
solutions already in the set. These initial solutions are then evaluated on the
actual objective function and the result is used to build the initial surrogate
model.% \citet{ZaeStoBar2014:ppsn} uses 10

@@

When we compare the behaviour of CEGO and uMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}).

 we 
: LOP syntéticas, todas en una gráfica.

Fig.~\ref{fig:loplib}: LOP reales: he puesto 3 de diferentes tamanos. Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)

kra32 es el unico del QAP en el que hay un comportaiento no random (considera distancia euclidea) Fig~\ref{fig:kra32}

PFSF, rec05, en un problema muy pequeno ninguno hace nada Fig.~\ref{fig:rec05}

Table~\ref{tab:results} shows

@@

\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/synthetic_LOP_combined}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
\end{figure}

\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
  \end{figure}

\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
\end{figure}
\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_qap_kra32_dat}
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
\end{figure}

>>>>

We use the implementation of GECO provided by the
authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the
authors of CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use and
the GA that optimizes the surrogate models uses a population size of 20,
crossover rate of 0.5, mutation rate of 1, tournament selection of size 2 and
probability of 0.9, interchange mutation (i.e., exchanging two randomly
selected elements of the permutation) and cycle crossover for permutations. The
authors of CEGO used a budget of $10^4$ evaluations using the surrogate-model
for each run of the GA, however, with such setting, each run of CEGO requires
more than a week of computation time, which was infeasible for us, hence, we
set a value of \textcolor{red}{$10^3$}.

Although it is never stated in the original paper, the implementation of CEGO
generates a set of initial solutions of size $\minit=10$ by means of a
max-min-distance sequential design: new solutions are added to the set
sequentially by maximizing the minimum distance to solutions already in the
set. These initial solutions are then evaluated on the actual objective
function and the result is used to build the initial surrogate
model.% \citet{ZaeStoBar2014:ppsn} uses 10

@@

%: LOP syntéticas, todas en una gráfica. 
\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/synthetic_LOP_combined}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
\end{figure}

When we compare the behaviour of CEGO and uMM on synthetic LOP instances
(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
uniformity of the instances. In particular, CEGO starts with a random sampling
of X solutions, with are quite poor as expected, followed by building and
updating the Kriging model based new evaluations, which leads to an impressive
improvement in less than 50 evaluations. However, progress is more or less
halted at this point and the model seems to have trouble generating better
solutions. Lower uniformity values (e.g., $\mu=5$) show worsening evaluations
despite the increase in solutions used to build the model.

On the other hand, uMM converges much more slowly than CEGO, only reaching the
same fitness values around evaluation 250. Yet, uMM is able to keep the rate of
improvement and completely overtake CEGO around evaluation 350. As in the case
of CEGO, lower uniformity values lead to worse performance, although the
difference is much smaller for uMM than it was for CEGO.

In summary, the Kriging model clearly helps CEGO to quickly identify good
solutions in very few evaluations. However, for some reason, CEGO is not able to
further improve those solutions given more data. This may due to the
model not being able to produce accurate predictions or the underlying
optimizer not being able to find improved solutions using the model.\MANUEL{perhaps we should measure the prediction error and see if it grows or decreases. That would answer this question.}

\newcommand{\supplement}{\citep{Supplementary}}

%: LOP reales: he puesto 3 de diferentes tamanos. Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
  \end{figure}

The fast initial convergence of CEGO for synthetic LOP instances does not seem
to translate to the instances available in LOPLIB~\citep{}. We show in
Fig.~\ref{fig:loplib} three examples with different permutation size, however,
the results are consistent for other LOPLIB instances (complete results
available as supplementary material \supplement). Due to extremely long
runtimes of CEGO and constraints in our computing system, we set a maximum
wall-clock time of 5 days for each run of CEGO. On these instances, the
behaviour of CEGO and uMM is much more similar (at least up to the termination
point of CEGO). It appears that our synthetic LOP instances have a fitness
landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
instances do not present such landscape. Comparing CEGO and uMM results, the
differences get smaller, in favour of uMM, with larger instance size. Our
conjecture is that building an accurate Bayesian model and searching for good
solutions on it becomes harder for larger fitness landscapes.

In the case of the PFSP, none of the algorithms are able to achieve a
  progress better than a random search would, as illustrated by the results
  (Fig.~\ref{fig:rec05}) obtained on \texttt{rec05}, which is the smallest PFSP
  instance considered here. The lack of any apparent convergence suggests that
  the models (both Bayesian and probabilistic) are not learning anything about
  the fitness landscape. Results for larger PFSP corroborate this conclusions
  (\supplement). Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
  that other distance measures are more suited than Kendall's-$\tau$ distance
  for the PFSP, the reported fitness differences between various distance
  metrics are small, hence, we conjecture that the same behaviour will be
  observed with other distance metrics.

\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
\end{figure}

The situation is very similar for the QAP, where both algorithms show the same
behaviour as in the PFSP for all instances tested, except for
\texttt{kra32}. As shown in Fig~\ref{fig:kra32}, in this instance both
algorithms are able to improve over the number of evaluations, although not
with ease. As in the synthetic LOP instances, we again observe that CEGO does
improve faster than UMM initially, yet, around evaluation 250, it appears to
get stuck and around evaluation 300, UMM overtakes it. We believe the reason is
the same as before, that is, the model is not able to adequately reflect the
ruggedness of the actual landscape and, hence, the best solutions found by the
GA searching the model are not improved solutions for the actual problem. Interestingly, \texttt{kra32} is rectangular
whereas \texttt{tho32} contains rectangular distances and \texttt{nug12} and \texttt{nug30} contain Manhattan distances of rectangular grids.\MANUEL{so they all have the same distance type!!!}

\begin{figure}
  \centering%
  \includegraphics[width=\textwidth]{img/fitness_real_qap_kra32_dat}
    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
\end{figure}

In terms of best fitness achieved on each run, and taking into account the
random behaviour observed in some instances and discussed above,
Table~\ref{tab:results} shows\MANUEL{TODO}

@@
